---
title: "The Sound Symbolism Experiment"
author: "Your Name"
date: "11/26/2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract
Replace this text by a very brief summary of the study and results

## Introduction 
Replace this text with a very brief motivation for the study and statement of hypotheses/predictions

## Materials and Methods
Replace this text with a brief description of the experimental setup with particular focus on the the dependent and independent variables

# Analysis
Write here a full specification of the statistical model here: what are the variables in the model? Fixed and random effects?

```{r, echo=FALSE}
# in this code block you include ...
# the call of libraries
# set working directory
# import of data set
# any other preprocessing if needed
# notice that echo = FALSE which means that this part of the analysis will not be visible in the report

```

```{r}
# In this code block you put the actual analysis code of your data

```

# Results
Replace this text with a short prose/numeric report of your results following APA standards 

## Discussion
Replace this text with a very brief prose summary and interpretation (no numbers) of the results and how they relate to the hypotheses
