---
title: "Experimental Methods I - portfolio 1"
author: "your name"
date: "9/26/2017"
output: word_document #change this if you want a different output format
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# libraries here

# set working directory here

# import data set here


```

# Portfolio 1: Datamining of the CogSci 2017 Personality Data

Use the tools you have acquired in the exercises to answer the following questions about the CogSci Personality Data. You answer each questions 'numerically' by using code in a Rmarkdown code block, 'visually' in a graph if that is requested in the assignment, and in 'prose', i.e. a short written interpretation that answers the question.
Remember to comment your code!

##1) Who, on average, is faster at the tongue twister task, right-handers or left-handers? 
(please provide numeric, visual, and verbal answer)
```{r}
# code goes here


# graph code goes here

```

Your written response goes here:


##2) Who prefers higher volumes of music, males or females?
(please provide numeric, visual, and verbal answer)
```{r}
# code goes here


# graph code goes here

```

Your written response goes here:

##3) Is there a systematic relation between occular dominance and the rabbit-duck illusion?
E.g. does people with right occular dominance see the duck more than the rabbit?
(please provide numeric and verbal answer. A graph is optional)

```{r}
# code goes here


# optional graph code goes here

```

Your written response goes here:

##4) Do we observe differences in variance (measured by standard deviation) in music consumption between those who have Danish as their native language and those who speaks a different native language?
(please provide numeric and verbal answer. Optionally you can also provide a graph)
```{r}
# code goes here


# optional graph code goes here

```

Your written response goes here:

##5) Calculate the average music consumption per day for those who can touch the floor and those who can't. Do you observe a relation between music consumption and preference for volume in the two groups (just from looking at the data and graph)?

(please provide numeric, visual, and verbal answer)
```{r}
# code goes here


# graph code goes here

```

Your written response goes here:
