---
title: 'Portfolio 3: The Reading Experiment'
author: "your name"
date: "10/31/2017"
output: word_document
---

## Portfolio 3: The Reading Experiment 

In this portfolio assignment, you will be conducting analyses of data from your reading experiment. The portfolio has two overall sections asking different questions: 1) A correlational section investigating what aspects of words predict reading time, and 2) an experimental section asking about the way that contextual expectancies affect our reading time by contrasting two conditions.

For each analyses, remember to make qualified choices of statistical test based on assumption testing.
Futhermore, for each analysis, provide a written report following APA conventions and a plot that illustrates the result  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# import your libraries and data in this code block

```

# 1) What properties of words predict word-by-word reading time?

Conduct minimum three correlational analyses where you investigate the relation between reading time and relevant predictors such as e.g. word length, frequencies, etc. 
 
```{r}
# code for your correlational study 1 
```

Report goes here: 

```{r}
# code for your correlational study 2 
```

Report goes here:

```{r}
# code for your correlational study 3 
```

Report goes here:

# 2) How does contextual expectations affect reading time?

Conduct a contrastive analysis of the two conditions in your reading experiment. Single out reading time of the words that differ between the two versions of your text/story and compare the means using a t-test. 

```{r}
# code for your t-test study  
```

Report goes here:
