---
title: "Autism"
author: "Sebastian Scott Engen"
date: "30 apr 2019"
output: html_document
---

```{r setup, include=FALSE}
library(pacman)
p_load(rethinking, tidyverse, brms, cowplot, dagitty, lavaan)
setwd("C:/Users/Dropbox/CogSci 4/Comp modeling/Portfolio-3-Autism")
# data loading
d <- read.csv("Assignment3Data.csv")
```

DAG

```{r DAG1}
g <- dagitty('dag {
    ASD_true [pos="0,2.5"]
    Q1 [pos="1,2"]
    Qn [pos="1,3"]
    VIQ [pos="2,4"]
    NVIQ [pos="2,3"]
    SIQ [pos="2,2"]
    ADOS [pos="2,1"]
    ASD_diagnosis [pos="3,2.5"]
    ASD_true -> Q1 -> VIQ -> ASD_diagnosis
    ASD_true -> Q1 -> NVIQ -> ASD_diagnosis
    ASD_true -> Q1 -> SIQ -> ASD_diagnosis
    ASD_true -> Q1 -> ADOS -> ASD_diagnosis
    ASD_true -> Qn -> VIQ -> ASD_diagnosis
    ASD_true -> Qn -> NVIQ -> ASD_diagnosis
    ASD_true -> Qn -> SIQ -> ASD_diagnosis
    ASD_true -> Qn -> ADOS -> ASD_diagnosis
}')
plot(g)
```


```{r DAG2}
g <- dagitty('dag {
    ASD_true [pos="0,2.5"]
    Q1 [pos="1,2"]
    Qn [pos="1,3"]
    VIQ [pos="2,4"]
    NVIQ [pos="2,3"]
    SIQ [pos="2,2"]
    ADOS [pos="2,1"]
    ASD_diagnosis [pos="3,2.5"]
    sampling [pos = "4, 2.5"]
    ASD_true -> Q1 -> VIQ -> ASD_diagnosis
    ASD_true -> Q1 -> NVIQ -> ASD_diagnosis
    ASD_true -> Q1 -> SIQ -> ASD_diagnosis
    ASD_true -> Q1 -> ADOS -> ASD_diagnosis
    ASD_true -> Qn -> VIQ -> ASD_diagnosis
    ASD_true -> Qn -> NVIQ -> ASD_diagnosis
    ASD_true -> Qn -> SIQ -> ASD_diagnosis
    ASD_true -> Qn -> ADOS -> ASD_diagnosis
    ASD_diagnosis -> sampling
}')
plot(g)
```

#A: 
#preprocessing, scaling

```{r }
#Filtering visit 1 & neurotypicals 
d1 <- d %>%
  filter(Visit == 1) %>%
  filter(ASD == 1)
#centering 
colname <- c("VerbalIQ", "NonVerbalIQ", "SocialIQ")
z_score <- function(var) {
  return( (var - mean(var) ) / sd(var) )
}
d2 <- d1 %>%
  mutate_at(vars(colname), z_score)
#standardizing 
d2$ADOS <- d2$ADOS / sd(d2$ADOS)
```


#B
Prior check. 
Same for all 3 models, so we do it for one. 

```{r}
# making a formula, automatic prior detection
m1 <- bf(ADOS ~ 1 + NonVerbalIQ)
#prior predictive check - Riccardo style 
prior = c(prior(normal(2.15, 1), class = "Intercept"), 
          prior(normal(0, 1), class = "b"), 
          prior(normal(0,1), class = "sigma"))
#model with = "only" (commented bc. of run-time)
m1_only <- brm(m1, family = gaussian, d2, 
                prior = prior, sample_prior = "only")  
#summarizing 
plot(m1_only)
pp_check(m1_only)
#simulating from prior - would be cool with lines. 
fit <- brm(m1, family = gaussian, d2, 
                prior = prior, sample_prior = TRUE)
p_samples <- prior_samples(fit)
dens(p_samples)
```

#C:
#Doing the 3 models 

```{r}
#model 1 - now weakly regularizing. 
nonverbal <- rethinking::map(
  alist(
    ADOS ~ dnorm(mu, sigma),
    mu <- a + b*NonVerbalIQ,
    a ~ dnorm(2.15, 1), #based on dnorm(10, 5) priors on natural scale. 
    b ~ dnorm(0, 1), #95% between -2, 2. 
    sigma ~ dnorm(0, 1) 
    ),
  data = d2)
#model 2
verbal <- rethinking::map(
  alist(
    ADOS ~ dnorm(mu, sigma),
    mu <- a + b*VerbalIQ,
    a ~ dnorm(2.15, 1), 
    b ~ dnorm(0, 1),
    sigma ~ dnorm(0, 1)
    ),
  data = d2)
#model 3
social <- rethinking::map(
  alist(
    ADOS ~ dnorm(mu, sigma),
    mu <- a + b*SocialIQ,
    a ~ dnorm(2.15, 1), 
    b ~ dnorm(0, 1),
    sigma ~ dnorm(0, 1)
    ),
  data = d2)
```

plots.

```{r plot precis of models}
plot(precis(nonverbal))
plot(precis(verbal))
plot(precis(social))
```

model quality for nonverbal

```{r}
NonVerbal.seq <- seq(from = -3, to = 3, by = 1)
sim.NonVerbal <- sim(nonverbal, data = list(NonVerbalIQ = NonVerbal.seq))
ADOS.PI <- apply(sim.NonVerbal, 2, PI, prob = 0.89)
mu <- link(nonverbal)
mu <- link(nonverbal, data = data.frame(NonVerbalIQ = NonVerbal.seq))
mu.mean <- apply(mu, 2, mean) 
mu.HPDI <- apply(mu, 2, HPDI, prob = 0.89)
```

```{r plot quality nonverbal}
plot(ADOS ~ NonVerbalIQ, d2, col = col.alpha(rangi2, 1))
lines(NonVerbal.seq, mu.mean)
# draw HPDI region for line
shade(mu.HPDI, NonVerbal.seq )
# draw PI region for simulated heights
shade(ADOS.PI, NonVerbal.seq ) 
```

assesing model for verbal

```{r}
Verbal.seq <- seq(from = -2, to = 2, by = 1)
sim.Verbal <- sim(verbal, data = list(VerbalIQ = Verbal.seq))
ADOS.PI <- apply(sim.Verbal, 2, PI, prob = 0.89)
mu <- link(verbal, data = data.frame(VerbalIQ = Verbal.seq))
mu.mean <- apply(mu, 2, mean) 
mu.HPDI <- apply(mu, 2, HPDI, prob = 0.89)
```

```{r model quality plot verbal}
plot(ADOS ~ VerbalIQ, d2, col = col.alpha(rangi2, 1))
lines(Verbal.seq, mu.mean)
# draw HPDI region for line
shade(mu.HPDI, Verbal.seq )
# draw PI region for simulated heights
shade(ADOS.PI, Verbal.seq ) 
```

assessing model for social

```{r}
Social.seq <- seq(from = -3, to = 3, by = 1)
sim.Social <- sim(social, data = list(SocialIQ = Social.seq))
ADOS.PI <- apply(sim.Social, 2, PI, prob = 0.89)
mu <- link(social, data = data.frame(SocialIQ = Social.seq))
mu.mean <- apply(mu, 2, mean) 
mu.HPDI <- apply(mu, 2, HPDI, prob = 0.89)
```

```{r model quality social}
plot(ADOS ~ SocialIQ, d2, col = col.alpha(rangi2, 1))
lines(Social.seq, mu.mean)
# draw HPDI region for line
shade(mu.HPDI, Social.seq )
# draw PI region for simulated heights
shade(ADOS.PI, Social.seq ) 
```

```{r normally distributed residuals}
res_normal <- function(model, outcome, predictor) {
  
  mu <- coef(model)['a'] + 
    coef(model)['b']*predictor
  
  m.resid <- outcome - mu
  dens(m.resid)
}
res_normal(nonverbal, d2$ADOS, d2$NonVerbalIQ)
res_normal(verbal, d2$ADOS, d2$VerbalIQ)
res_normal(social, d2$ADOS, d2$SocialIQ)
```

TASK 2: 
#A different portions of variance?

```{r portions of variance}
#corr-check 
d2cor <- d2[6:8] #weird ones 
cor(d2cor) #social & nonverbal are not that correlated. 
```

```{r residuals}
### models predictor ~ predictor
r_nv <- rethinking::map(
  alist(
    NonVerbalIQ ~ dnorm(mu, sigma),
    mu <- a + b1*VerbalIQ + b2*SocialIQ,
    a ~ dnorm(0, 1), 
    b1 ~ dnorm(0, 1),
    b2 ~ dnorm(0, 1),
    sigma ~ dnorm(0, 1)
    ),
  data = d2)
r_s <- rethinking::map(
  alist(
    SocialIQ ~ dnorm(mu, sigma),
    mu <- a + b1*VerbalIQ + b2*NonVerbalIQ,
    a ~ dnorm(0, 1), 
    b1 ~ dnorm(0, 1),
    b2 ~ dnorm(0, 1),
    sigma ~ dnorm(0, 1)
    ),
  data = d2)
r_v <- rethinking::map(
  alist(
    VerbalIQ ~ dnorm(mu, sigma),
    mu <- a + b1*NonVerbalIQ + b2*SocialIQ,
    a ~ dnorm(0, 1), 
    b1 ~ dnorm(0, 1),
    b2 ~ dnorm(0, 1),
    sigma ~ dnorm(0, 1)
    ),
  data = d2)
# RESIDUALS AGAINST ADOS
## funciton
resi <- function(model, outcome, predictor1, predictor2) {
  
  mu <- coef(model)['a'] + 
    coef(model)['b1']*predictor1 + coef(model)['b2']*predictor2
  
  m.resid <- outcome - mu #outcome - what we predicted. 
  return(m.resid)
}
# unique information in Verbal IQ
d2 %>%
  mutate(res = resi(r_v, d2$VerbalIQ, d2$NonVerbalIQ, d2$SocialIQ)) %>%
  ggplot(aes(res, ADOS)) +
  geom_point() +
  geom_smooth(method = "lm")
# unique information in NonVerbal IQ
d2 %>%
  mutate(res = resi(r_nv, d2$NonVerbalIQ, d2$VerbalIQ, d2$SocialIQ)) %>%
  ggplot(aes(res, ADOS)) +
  geom_point() +
  geom_smooth(method = "lm")
# unique information in Social IQ
d2 %>%
  mutate(res = resi(r_s, d2$SocialIQ, d2$VerbalIQ, d2$NonVerbalIQ)) %>%
  ggplot(aes(res, ADOS)) +
  geom_point() +
  geom_smooth(method = "lm")
```


```{r multiple predictor models}
set.seed(711)
three_pred <- rethinking::map(
alist(
ADOS ~ dnorm(mu, sigma),
mu <- Intercept + b_NonVIQ*NonVerbalIQ + b_VIQ * VerbalIQ + b_SIQ * SocialIQ,
Intercept ~ dnorm(2.15, 1),
b_NonVIQ ~ dnorm(0, 1),
b_VIQ ~ dnorm(0, 1), 
b_SIQ ~ dnorm(0, 1),
sigma ~ dnorm(0, 1)
),
data = d2)
set.seed(911)
two_pred <- rethinking::map(
alist(
ADOS ~ dnorm(mu, sigma),
mu <- Intercept + b_NonVIQ*NonVerbalIQ + b_SIQ * SocialIQ,
Intercept ~ dnorm(2.15, 1),
b_NonVIQ ~ dnorm(0, 1),
b_SIQ ~ dnorm(0, 1),
sigma ~ dnorm(0, 1)
),
data = d2)
```


```{r precis of 5 models}
precis(verbal)
precis(nonverbal)
precis(social)
precis(three_pred, cor = TRUE)
precis(two_pred, cor = TRUE)
```

NB: takes some time to run next chunk. 

```{r WAIC of 5 models}
set.seed(1337)
model_comp <- compare(verbal, nonverbal, social, three_pred, two_pred, n = 10^4)
model_comp  
```

2C # Assessing quality of predictors. 
Model averaged posterior predictive distribution

```{r MAPPD non-verbal}
non.seq <- seq(from=-2.5,to=3,length.out=30)
d.predict <- list(
ADOS = rep(0,30), # empty outcome 
NonVerbalIQ = non.seq, # sequence of social IQ
SocialIQ = rep(0,30), # average of non-verbal
VerbalIQ = rep(0,30) #average of verbal 
)
pred.three <- link( three_pred , data=d.predict )
mu <- apply( pred.three , 2 , mean )
mu.PI <- apply( pred.three , 2 , PI )
# plot it all
plot( ADOS ~ NonVerbalIQ , d2 , col=rangi2 )
lines( non.seq , mu , lty=2 )
lines( non.seq , mu.PI[1,] , lty=2 )
lines( non.seq , mu.PI[2,] , lty=2 )
#ensemble
IQ.ensemble <- ensemble(verbal, nonverbal, social, three_pred, two_pred, data = d.predict)
mu <- apply( IQ.ensemble$link , 2 , mean )
mu.PI <- apply( IQ.ensemble$link , 2 , PI )
lines( non.seq , mu )
shade( mu.PI , non.seq )
```

2C # Assessing quality of predictors. 
This block is for verbal 

```{r MAPPD verbal}
# mean & spread. 
summary(d2$SocialIQ)
summary(d2$NonVerbalIQ) #0
summary(d2$VerbalIQ) #0
verb.seq <- seq(from=-1.5,to=2,length.out=30)
d.predict <- list(
ADOS = rep(0,30), # empty outcome 
VerbalIQ = verb.seq, # sequence of social IQ
SocialIQ = rep(0,30), # average of non-verbal
NonVerbalIQ = rep(0,30) #average of verbal 
)
pred.three <- link( three_pred , data=d.predict )
mu <- apply( pred.three , 2 , mean )
mu.PI <- apply( pred.three , 2 , PI )
# plot it all
plot( ADOS ~ VerbalIQ , d2 , col=rangi2 )
lines( verb.seq , mu , lty=2 )
lines( verb.seq , mu.PI[1,] , lty=2 )
lines( verb.seq , mu.PI[2,] , lty=2 )
#ensemble
IQ.ensemble <- ensemble(verbal, nonverbal, social, three_pred, two_pred, data = d.predict)
mu <- apply( IQ.ensemble$link , 2 , mean )
mu.PI <- apply( IQ.ensemble$link , 2 , PI )
lines( verb.seq , mu )
shade( mu.PI , verb.seq )
```

2C # Assessing quality of predictors. 
This block is for social

```{r MAPPD social}
# mean & spread. 
summary(d2$SocialIQ)
summary(d2$NonVerbalIQ) #0
summary(d2$VerbalIQ) #0
soc.seq <- seq(from=-1.5,to=2.5,length.out=30)
d.predict <- list(
ADOS = rep(0,30), # empty outcome 
SocialIQ = soc.seq, # sequence of social IQ
VerbalIQ = rep(0,30), # average of non-verbal
NonVerbalIQ = rep(0,30) #average of verbal 
)
pred.three <- link( three_pred , data=d.predict )
mu <- apply( pred.three , 2 , mean )
mu.PI <- apply( pred.three , 2 , PI )
# plot it all
plot( ADOS ~ SocialIQ , d2 , col=rangi2 )
lines( soc.seq , mu , lty=2 )
lines( soc.seq , mu.PI[1,] , lty=2 )
lines( soc.seq , mu.PI[2,] , lty=2 )
#ensemble
IQ.ensemble <- ensemble(verbal, social, nonverbal, three_pred, two_pred, data = d.predict)
mu <- apply( IQ.ensemble$link , 2 , mean )
mu.PI <- apply( IQ.ensemble$link , 2 , PI )
lines( soc.seq , mu )
shade( mu.PI , soc.seq )
```

Task 3 prep 

WITH TD AND ASD
```{r non-sequitor, include = F}
#MAKE DATAFRAME
d3 <- d %>%
  filter(Visit == 1)
colname <- c("VerbalIQ", "NonVerbalIQ", "SocialIQ")
d4 <- d3[complete.cases(d3[6:8]) , ]
d4 <- d4 %>%
  mutate_at(vars(colname), z_score) %>%
  mutate(ADOS = ADOS / sd(ADOS))
```

Task 3
Interaction (only non-verbal as main effect).
Just kept in there for now as a baseline (how to do it).
Also, this works for sure.

```{r interaction}
set.seed(911)
library(rethinking)
#weakly regularizing priors & linear interaction. 
nonv_int <- rethinking::map( 
alist(
ADOS ~ dnorm( mu , sigma ) ,
mu <- intercept + gamma*NonVerbalIQ + bgroup*ASD ,
gamma <- bnon + binteraction*ASD ,
intercept ~ dnorm( 2.15 , 1 ) , #intercept
bnon ~ dnorm( 0 , 1 ) , #non verbal 
bgroup ~ dnorm( 0 , 1 ) , #dummy/index
binteraction ~ dnorm( 0 , 1 ) , #by group
sigma ~ dnorm( 0 , 1 ) #sigma
) ,
data=d4,
method="Nelder-Mead" ,
control=list(maxit=1e4)
)
precis(nonv_int)
plot(precis(nonv_int))
#plotting
IQ.seq <- seq(from=-3,to=3,by=0.25) 
mu.ASD <- link( nonv_int, data=data.frame(ASD=1,NonVerbalIQ=IQ.seq) )
mu.ASD.mean <- apply( mu.ASD , 2 , mean )
mu.ASD.PI <- apply( mu.ASD , 2 , PI , prob=0.97 )
mu.TD <- link( nonv_int , data=data.frame(ASD=0,NonVerbalIQ=IQ.seq) )
mu.TD.mean <- apply( mu.TD , 2 , mean )
mu.TD.PI <- apply( mu.TD , 2 , PI , prob=0.97 )
# regression by ADOS.
d.A1 <- d4[d4$ASD==1,]
plot( ADOS ~ NonVerbalIQ , data=d.A1 ,
col=rangi2 , ylab="ADOS" ,
xlab="NonVerbal" )
mtext( "ASD" , 3 )
lines( IQ.seq , mu.ASD.mean , col=rangi2 )
shade( mu.ASD.PI , IQ.seq , col=col.alpha(rangi2,0.3) )
# regression by TD. 
d.A0 <- d4[d4$ASD==0,]
plot( ADOS ~ NonVerbalIQ , data=d.A0 ,
col="black" , ylab="ADOS" ,
xlab="NonVerbalIQ" )
mtext( "TD" , 3 )
lines( IQ.seq , mu.TD.mean)
shade( mu.TD.PI , IQ.seq)
#computing stuff
post <- extract.samples( nonv_int ) 
gamma.ASD <- post$bnon + post$binteraction*1
gamma.TD <- post$bnon + post$binteraction*0
mean( gamma.ASD) #effect of NON-verbal on ASD 
mean( gamma.TD ) #effect of NON-verbal on TD
dens( gamma.ASD , xlim=c(-0.5,0.5) , ylim=c(0,8) , 
xlab="gamma" , col=rangi2 )
dens( gamma.TD , add=TRUE )
```

Task 3 Interaction.
Interaction with all of the variables. 

```{r}
mallinteract <- rethinking::map(
  alist(
    ADOS ~ dnorm(mu, sigma),
    mu <- intercept + gamma1*NonVerbalIQ + bgroup*ASD + 
          gamma2*VerbalIQ + bgroup*ASD + gamma3*SocialIQ + bgroup*ASD,
    gamma1 <- bnon + binteraction1*ASD, #nonverbal
    gamma2 <- bverb + binteraction2*ASD, #verbal
    gamma3 <- bsoc + binteraction3*ASD, #social
    intercept ~ dnorm(2.15, 1),
    bnon ~ dnorm(0, 1),
    bverb ~ dnorm(0, 1),
    bsoc ~ dnorm(0,1),
    bgroup ~ dnorm(0,1),
    binteraction1 ~ dnorm(0,1),
    binteraction2 ~ dnorm(0,1),
    binteraction3 ~ dnorm(0,1),
    sigma ~ dnorm(0,1)
  ),
  data = d4,
  method="Nelder-Mead" ,
control=list(maxit=1e4)
)
precis(mallinteract)
compare(mallinteract, nonv_int) #multiple interaction best. 
plot(precis(mallinteract))
#extracting samples nonverbal
post <- extract.samples( mallinteract ) 
gamma.ASD.non <- post$bnon + post$binteraction1*1
gamma.TD.non <- post$bnon + post$binteraction1*0
mean(gamma.ASD.non) #more negative. 
mean(gamma.TD.non) #less negative. 
#extracting samples verbal
gamma.ASD.verb <- post$bverb + post$binteraction2*1
gamma.TD.verb <- post$bverb + post$binteraction2*0
mean(gamma.ASD.verb)
mean(gamma.TD.verb)
#extracting samples social
gamma.ASD.soc <- post$bsoc + post$binteraction3*1
gamma.TD.soc <- post$bsoc + post$binteraction3*0
mean(gamma.ASD.soc)
mean(gamma.TD.soc)
```

TASK 3:
visually inspecting whether IQ scores make sense for 
ASD and TD children respectively. 
Specifically checking for interaction effect (which is there..)

```{r}
#Is there an interaction 
d5 <- rbind(d2, d4)
d5$ASD <- as.factor(d5$ASD)
#checking with plots 
library(tidyverse)
library(ggunchained)
p1 <- ggplot(d5, aes(NonVerbalIQ, ADOS, color = ASD)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "ADOS score by non-verbal IQ for ASD and TD") 
p2 <- ggplot(d5, aes(VerbalIQ, ADOS, color = ASD)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "ADOS score by verbal IQ for ASD and TD") 
p3 <- ggplot(d5, aes(SocialIQ, ADOS, color = ASD)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "ADOS score by social IQ for ASD and TD")
```


```{r cowplot}
plot_grid(p1, p2, p3)
```

more variation? 
delete Magnus if there is no need for it. 

```{r}
library(ggridges)
#some ASD very high social IQ? high density towards lower bound for ASD. 
p11 <- ggplot(d5, aes(SocialIQ, ASD, fill = ASD)) + 
  geom_density_ridges()
#not as many high-performing ASD as TD. 
p12 <- ggplot(d5, aes(VerbalIQ, ASD, fill = ASD)) +
  geom_density_ridges() 
#crazy tail here --> some very high non-verbal IQ
p13 <- ggplot(d5, aes(NonVerbalIQ, ASD, fill = ASD)) +
  geom_density_ridges() 
#
plot_grid(p11, p12, p13)
```

Question 5 

ADOS SE = 1.
still somewhat messy, but it's understandable. 

```{r no outputs just new model}
#SE of ados?
d7 <- d1 %>%
  mutate(ADOS.SE = 1/sd(ADOS))
d8 <- cbind(d2, d7[[10]])
#renaming. 
d8 <- d8 %>%
  rename("ADOS_SE" = "d7[[10]]")
dlist <- list( 
div_obs=d8$ADOS,
div_sd=d8$ADOS_SE, #how do we convert?
G=d8$NonVerbalIQ #non-verbal
)
m14.1 <- rethinking::map2stan(
alist(
div_est ~ dnorm(mu,sigma),
mu <- a + b*G,
div_obs ~ dnorm(div_est,div_sd),
a ~ dnorm(3,1), #intercept
b ~ dnorm(-0.5,0.5),
sigma ~ dnorm(1,0.5) #sigma
) , 
data=dlist , 
start=list(div_est=dlist$div_obs) , 
WAIC=FALSE , iter=5000 , warmup=1000 , chains=2 , cores=2 ,
control=list(adapt_delta=0.95) )
#model 1
nonverbal <- rethinking::map(
  alist(
    ADOS ~ dnorm(mu, sigma),
    mu <- a + b*NonVerbalIQ,
    a ~ dnorm(3, 1), 
    b ~ dnorm(-0.5, 0.5),
    sigma ~ dnorm(1, 0.5)
    ),
  data = d2)
```

preparing to include the measurement error on IQ. 

```{r more messy stuff without }
#column for it 
d9 <- d1 %>%
  mutate(nonIQ_SE = 2.12 / sd(NonVerbalIQ))
d8 <- cbind(d8, d9$nonIQ_SE)
d8 <- d8 %>%
  rename("nonIQ_SE" = "d9$nonIQ_SE")
```


```{r uncertainty in both predictor and outcome}
dlist <- list( 
ADOS_obs=d8$ADOS,
ADOS_sd=d8$ADOS_SE, #how do we convert?
non_obs=d8$NonVerbalIQ, #non-verbal
non_sd=d8$nonIQ_SE
)
#map2stand
m14.2 <- map2stan(
alist(
ADOS_est ~ dnorm(mu,sigma),
mu <- a + b*non_est[i],
ADOS_obs ~ dnorm(ADOS_est,ADOS_sd),
non_obs ~ dnorm(non_est,non_sd),
a ~ dnorm(2.15,1), #intercept
b ~ dnorm(0,1),
sigma ~ dnorm(0,1) 
) ,
data=dlist ,
start=list(ADOS_est=dlist$ADOS_obs, non_est=dlist$non_obs) ,
WAIC=FALSE , iter=5000 , warmup=1000 , chains=3 , cores=3 ,
control=list(adapt_delta=0.95) )
```

```{r comparing precis all models}
precis(m14.1) #only modeling measurement error on outcome. 
precis(m14.2) #biggest sd, but same beta as original 
precis(nonverbal) #original
```

```{r precis plot all models}
plot(precis(m14.1)) 
plot(precis(m14.2)) 
plot(precis(nonverbal))
```

