---
title: "Computational Modeling - Week 3 - Assignment 2 - Part 1"
author: "Riccardo Fusaroli"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## In this assignment we learn how to assess rates from a binomial distribution, using the case of assessing your teachers' knowledge of CogSci

### First part

You want to assess your teachers' knowledge of cognitive science. "These guys are a bunch of drama(turgist) queens, mindless philosophers, chattering communication people and Russian spies. Do they really know CogSci?", you think.

To keep things simple (your teachers should not be faced with too complicated things):
- You created a pool of equally challenging questions on CogSci
- Each question can be answered correctly or not (we don't allow partially correct answers, to make our life simpler).
- Knowledge of CogSci can be measured on a scale from 0 (negative knowledge, all answers wrong) through 0.5 (random chance) to 1 (awesome CogSci superpowers)

This is the data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
- Mikkel: 66 correct answers out of 132 questions

Questions:

1. What's Riccardo's estimated knowledge of CogSci? What is the probability he knows more than chance (0.5) [try figuring this out. if you can't peek into chapters 3.1 and 3.2 and/or the slides]?
- First implement a grid approximation (hint check paragraph 2.4.1!) with a uniform prior, calculate the posterior and plot the results
```{r}
#Ricardo
p_grid <- seq( from=0 , to=1 , length.out=1000 ) #this is some strong overfitting; in reality, we have only 6 data-points (questions) to draw our knowledge from
prior <- rep( 1 , 1000 ) #flat prior, thus all 1s
likelihood <- dbinom( 3 , size=6 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)

plot(p_grid, posterior, xlab="proportion of correct answers")


```

- Then implement a quadratic approximation (hint check paragraph 2.4.2!).
```{r}
install.packages(c("coda","mvtnorm","devtools"))
devtools::install_github("rmcelreath/rethinking")

library(devtools)
library(rethinking)
globe.qa <- map(alist(q ~ dbinom(6,p) ,  # binomial likelihood - total n questions
                      p ~ dunif(0,1)     # uniform prior
                      ), data=list(q=3) ) #correct n questions, riccardo
# display summary of quadratic approximation
precis( globe.qa )
"""
Mean StdDev 5.5% 94.5%
p  0.5    0.2 0.17  0.83
#Mean, Std and confidence intervals
"""
```
- N.B. for the rest of the exercise just keep using the grid approximation (we'll move to quadratic approximations in two classes)


2. Estimate all the teachers' knowledge of CogSci. Who's best? Use grid approximation. Comment on the posteriors of Riccardo and Mikkel (the 50%s; but one with more data than the other.
```{r 2.}
#Mikkel
p_grid <- seq( from=0 , to=1 , length.out=1000 ) #this is some strong overfitting; in reality, we have only 6 data-points (questions) to draw our knowledge from
prior <- rep( 1 , 1000 ) #flat prior, thus all 1s
likelihood <- dbinom( 66 , size=132 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)

plot(p_grid, posterior, xlab="proportion of correct answers")


#Josh
p_grid <- seq( from=0 , to=1 , length.out=10000 ) #this is some strong overfitting; in reality, we have only 6 data-points (questions) to draw our knowledge from
prior <- rep( 1 , 10000 ) #flat prior, thus all 1s
likelihood <- dbinom( 160 , size=198 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)

plot(p_grid, posterior, xlab="proportion of correct answers")

#Tylen
p_grid <- seq( from=0 , to=1 , length.out=1000 ) #this is some strong overfitting; in reality, we have only 6 data-points (questions) to draw our knowledge from
prior <- rep( 1 , 1000 ) #flat prior, thus all 1s
likelihood <- dbinom( 2 , size=2 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)

plot(p_grid, posterior, xlab="proportion of correct answers")
plot(p_grid, prior, xlab="proportion of correct answers") # Flat prior - No expectations
```

2a. Produce plots of the prior, and posterior for each teacher.
```{r 2a.}
#rethinking

```


3. Change the prior. Given your teachers have all CogSci jobs, you should start with a higher appreciation of their knowledge: the prior is a normal distribution with a mean of 0.8 and a standard deviation of 0.2. Do the results change (and if so how)?
```{r}
#Ricardo
p_grid <- seq( from=0 , to=1 , length.out=100 )
prior <- dnorm(p_grid,0.8,0.2) #centered around mean of 0.8, std of 0.2
likelihood <- dbinom( 3 , size=6 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)

dens = 1000
plot( p_grid , posterior , type="b" ,
    xlab="probability of cogsci knowledge" , ylab="posterior probability" )
mtext( "20 points" )
Data=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood)
ggplot(Data,aes(grid,posterior, color='red'))+  geom_point()+theme_classic()+  geom_line(aes(grid,prior/dens*10),color='cyan')+  xlab("Knowledge of CogSci")+ ylab("plausability") + theme(legend.position = "none")
#skewing the data towards more correct answers (riccardos answers at 0.5 are getting an upward-nudge)
# We do not only rely on the data, but also on our prior knowledge of Ricardo as a teacher

#Mikkel
p_grid <- seq( from=0 , to=1 , length.out=1000 ) #this is some strong overfitting; in reality, we have only 6 data-points (questions) to draw our knowledge from
prior <- dnorm(p_grid,0.8,0.2) #centered around mean of 0.8, std of 0.2
likelihood <- dbinom( 66 , size=132 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)

dens = 1000
plot( p_grid , posterior , type="b" ,
    xlab="probability of cogsci knowledge" , ylab="posterior probability" )
mtext( "20 points" )
Data=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood)
ggplot(Data,aes(grid,posterior, color='red'))+  geom_point()+theme_classic()+  geom_line(aes(grid,prior/dens*3),color='cyan')+  xlab("Knowledge of CogSci")+ ylab("plausability") + theme(legend.position = "none")


#Josh
p_grid <- seq( from=0 , to=1 , length.out=10000 ) #this is some strong overfitting; in reality, we have only 6 data-points (questions) to draw our knowledge from
prior <- dnorm(p_grid,0.8,0.2) #centered around mean of 0.8, std of 0.2
likelihood <- dbinom( 160 , size=198 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)

dens = 1000
plot( p_grid , posterior , type="b" ,
    xlab="probability of cogsci knowledge" , ylab="posterior probability" )
mtext( "20 points" )
Data=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood)
ggplot(Data,aes(grid,posterior, color='red'))+  geom_point()+theme_classic()+  geom_line(aes(grid,prior/dens/3),color='cyan')+  xlab("Knowledge of CogSci")+ ylab("plausability") + theme(legend.position = "none")

#Tylen
p_grid <- seq( from=0 , to=1 , length.out=1000 ) #this is some strong overfitting; in reality, we have only 6 data-points (questions) to draw our knowledge from
prior <- dnorm(p_grid,0.8,0.2) #centered around mean of 0.8, std of 0.2
likelihood <- dbinom( 2 , size=2 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)

dens = 1000
plot( p_grid , posterior , type="b" ,
    xlab="probability of cogsci knowledge" , ylab="posterior probability" )
mtext( "20 points" )
Data=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood)
ggplot(Data,aes(grid,posterior, color='red'))+  geom_point()+theme_classic()+  geom_line(aes(grid,prior/dens*1.3),color='cyan')+  xlab("Knowledge of CogSci")+ ylab("plausability") + theme(legend.position = "none")
```



3a. Produce plots of the prior and posterior for each teacher.
```{r}
plot(p_grid, prior) #very positively skewed
plot(p_grid, posterior) #riccardos bad results are drawing the distribution towards random-guessing (0.5, that is)

dens = 1000
plot( p_grid , posterior , type="b" ,
    xlab="probability of cogsci knowledge" , ylab="posterior probability" )
mtext( "20 points" )
Data=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood)
ggplot(Data,aes(grid,posterior, color='red'))+  geom_point()+theme_classic()+  geom_line(aes(grid,prior/dens*10),color='cyan')+  xlab("Knowledge of CogSci")+ ylab("plausability") + theme(legend.position = "none")

```

4. You go back to your teachers and collect more data (multiply the previous numbers by 100). Calculate their knowledge with both a uniform prior and a normal prior with a mean of 0.8 and a standard deviation of 0.2. Do you still see a difference between the results? Why?
```{r}
#Ricardo
p_grid <- seq( from=0 , to=1 , length.out=1000 ) #this is some strong overfitting; in reality, we have only 6 data-points (questions) to draw our knowledge from
prior <- dnorm(p_grid,0.8,0.2) #flat prior, thus all 1s
likelihood <- dbinom( 300 , size=600 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)

plot(p_grid, posterior, xlab="proportion of correct answers")

dens = 1000
plot( p_grid , posterior , type="b" ,
    xlab="probability of cogsci knowledge" , ylab="posterior probability" )
mtext( "20 points" )
Data=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood)
ggplot(Data,aes(grid,posterior, color='red'))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='cyan')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")




#Mikkel
p_grid <- seq( from=0 , to=1 , length.out=1000 ) #this is some strong overfitting; in reality, we have only 6 data-points (questions) to draw our knowledge from
prior <- dnorm(p_grid,0.8,0.2) #flat prior, thus all 1s
likelihood <- dbinom( 6600 , size=13200 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)

plot(p_grid, posterior, xlab="proportion of correct answers")

dens = 1000
plot( p_grid , posterior , type="b" ,
    xlab="probability of cogsci knowledge" , ylab="posterior probability" )
mtext( "20 points" )
Data=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood)
ggplot(Data,aes(grid,posterior, color='red'))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='cyan')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")




#Josh
p_grid <- seq( from=0 , to=1 , length.out=10000 ) #this is some strong overfitting; in reality, we have only 6 data-points (questions) to draw our knowledge from
prior <- dnorm(p_grid,0.8,0.2) #flat prior, thus all 1s
likelihood <- dbinom( 16000 , size=19800 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)

plot(p_grid, posterior, xlab="proportion of correct answers")

dens = 1000
plot( p_grid , posterior , type="b" ,
    xlab="probability of cogsci knowledge" , ylab="posterior probability" )
mtext( "20 points" )
Data=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood)
ggplot(Data,aes(grid,posterior, color='red'))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='cyan')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")



#Tylen
p_grid <- seq( from=0 , to=1 , length.out=1000 ) #this is some strong overfitting; in reality, we have only 6 data-points (questions) to draw our knowledge from
prior <- dnorm(p_grid,0.8,0.2) #flat prior, thus all 1s
likelihood <- dbinom( 200 , size=200 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)

plot(p_grid, posterior, xlab="proportion of correct answers")
plot(p_grid, prior, xlab="proportion of correct answers") # Flat prior - No expectations

dens = 1000
plot( p_grid , posterior , type="b" ,
    xlab="probability of cogsci knowledge" , ylab="posterior probability" )
mtext( "20 points" )
Data=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood)
ggplot(Data,aes(grid,posterior, color='red'))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='cyan')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")
```


5. Imagine you're a skeptic and think your teachers do not know anything about CogSci, given the content of their classes. How would you operationalize that belief?
```{r}
#Ricardo
p_grid <- seq( from=0 , to=1 , length.out=1000 ) #this is some strong overfitting; in reality, we have only 6 data-points (questions) to draw our knowledge from
prior <- dnorm(p_grid,0.5,0.2) # normal chance
likelihood <- dbinom( 3 , size=6 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)

plot(p_grid, posterior, xlab="proportion of correct answers")

dens = 1000
plot( p_grid , posterior , type="b" ,
    xlab="probability of cogsci knowledge" , ylab="posterior probability" )
mtext( "20 points" )
Data=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood)
ggplot(Data,aes(grid,posterior, color='red'))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='cyan')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")




#Mikkel
p_grid <- seq( from=0 , to=1 , length.out=1000 ) #this is some strong overfitting; in reality, we have only 6 data-points (questions) to draw our knowledge from
prior <- dnorm(p_grid,0.5,0.2) # normal chance
likelihood <- dbinom( 66 , size=132 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)

plot(p_grid, posterior, xlab="proportion of correct answers")

dens = 1000
plot( p_grid , posterior , type="b" ,
    xlab="probability of cogsci knowledge" , ylab="posterior probability" )
mtext( "20 points" )
Data=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood)
ggplot(Data,aes(grid,posterior, color='red'))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='cyan')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")




#Josh
p_grid <- seq( from=0 , to=1 , length.out=10000 ) #this is some strong overfitting; in reality, we have only 6 data-points (questions) to draw our knowledge from
prior <- dnorm(p_grid,0.5,0.2) # normal chance
likelihood <- dbinom( 160 , size=198 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)

plot(p_grid, posterior, xlab="proportion of correct answers")

dens = 1000
plot( p_grid , posterior , type="b" ,
    xlab="probability of cogsci knowledge" , ylab="posterior probability" )
mtext( "20 points" )
Data=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood)
ggplot(Data,aes(grid,posterior, color='red'))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='cyan')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")



#Tylen
p_grid <- seq( from=0 , to=1 , length.out=1000 ) #this is some strong overfitting; in reality, we have only 6 data-points (questions) to draw our knowledge from
prior <- dnorm(p_grid,0.5,0.2) # normal chance
likelihood <- dbinom( 2 , size=2 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)

plot(p_grid, posterior, xlab="proportion of correct answers")
plot(p_grid, prior, xlab="proportion of correct answers") # Flat prior - No expectations

dens = 1000
plot( p_grid , posterior , type="b" ,
    xlab="probability of cogsci knowledge" , ylab="posterior probability" )
mtext( "20 points" )
Data=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood)
ggplot(Data,aes(grid,posterior, color='red'))+  geom_point()+geom_line()+theme_classic()+  geom_line(aes(grid,prior/dens),color='cyan')+  xlab("Knowledge of CogSci")+ ylab("posterior probability")

#Tylen is heavily influenced by our priors since there's only two datapoints to go on
```



6. Optional question: Can you estimate the difference between Riccardo's estimated knowledge and that of each of the other teachers? Would you deem it credible (that is, would you believe that it is actually different)?

7. Bonus knowledge: all the stuff we have done can be implemented in a lme4-like fashion using the brms package. Here is an example.
```{r}
library(brms)
d <- data.frame(
  Correct=c(3,2,160,66),
  Questions=c(6,2,198,132),
  Teacher=c("RF","KT","JS","MW"))

FlatModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("uniform(0,1)", class = "Intercept"),family=binomial)
plot(FlatModel)
PositiveModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.8,0.2)", class = "Intercept"),family=binomial)
plot(PositiveModel)
SkepticalModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.5,0.01)", class = "Intercept"),family=binomial)
plot(SkepticalModel)
```

If you dare, try to tweak the data and model to test two hypotheses:
- Is Kristian different from Josh?
- Is Josh different from chance?



